Modeling Second-Language Learning from a Psychological Perspective

### 2018-03-27 doc/naaclhlt2018-latex/paper.tex
add discussion

>>>>

\section{Discussion}

When approaching the Duolingo SLAM task, we hoped to leverage psychological
insights in building our model. We found that in some cases, such as the
age of acquisition, this was helpful. In general, though, our model gained its
power not from hand-crafted features but from applying a powerful inference
technique (gradient boosted trees) to raw input about user and word IDs and
exercise features.

There are multiple reasons for the limited applicability of psychology to this
competition. First, computational psychological models, while useful, are not
well-suited to generating highly accurate predictions from large data sets.
Because they are designed not for prediction but for explanation, they tend to
use a small number of input variables and to allow those variables to interact
in limited ways. In contrast, gradient boosted trees, as well as other
cutting-edge techniques like deep learning. can extract high-level
interactions among hundreds of features. Though they are highly opaque, require
a lot of data, and are not amenable to explanation, these models excel at prediction.

Second, it is possible that our ability to use theories of learning, including
ideas about massed and spaced practice, was disrupted by the fact that the data
may have been adaptively created using these very principles
\cite{settles2016trainable}. If Duolingo adaptively sequenced the spacing of
trials based on past errors, then the relationship between future errors and
past spacing may have differed from that found in the psychological literature
\cite{Cepeda:2006aa}.

Finally, a task in which more true generalization was required might have
allowed psychologically inspired features to perform more competitively. In this
task, there is a large amount of labeled training data for every user and for
most words. This allow simple ID-based features to work. However, for a
newly-encountered user or word, an ID is useless. Theory-driven features, in
contrast, can generalize to new settings. Thus, if we were asked to generalize to
a completely new language such as German, many parts of our model would falter
but word frequency, age of acquisition, and Levenshtein distance to
first-language translation would still likely be predictive.
